---
title: "Tracking Habits: An analysis on respondents screen time"
description: |
  Analyzing data using a combination of Google Forms, Google Sheets and R Studio.
author:
  - name: Natasha Munn 
date: 07-07-2024
output:
  distill::distill_article:
    self_contained: false
    code_folding: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I love a good Google Form. Therefore, it would be fun to gather data using Forms, analyse it on R Studio and see what we can learn!

Today, I am investigating device use and screen time to explore the differences in my respondent's screen time habits (*i.e., time spent, devices most used, applications most used*).

My first step was to create my form. It is vital to consider data ethics when collecting information, so I communicated my purpose to my respondents in my form description. I also clearly stated who was collecting the data, who would have access to the data, and what the data would be used for. I also communicated that their responses would be anonymous and that the only additional identifiable information would be collected would be the date and time they submitted the form. 

I phrased my questions for the respondent only to measure their screen time habits of the last week or ask them to answer the question based on their current stance. I asked the following questions:

1. Which age category best describes you?
2. Which of the following devices do you use on a regular basis? (at least once a week)
3. Out of the following list of devices, which one did you use the most in the last week?
4. In the last week, how many hours of screen time did you spend in total? 
5. Which platform has taken up the most of your screen time in the last week? 
6. Are you currently employed
7. (if yes to Q6) Is working on a device with a screen an essential part of your job on a weekly basis?
8. In the last week, how many nights have you used a device with a screen directly before bed?

Now, let's start investigating. First, we need to load **tidyverse**, as well as the **magick** package for a special surprise later on:)

```{r}
library(tidyverse)
library(magick)
```

I connected my form to a Google Sheet, which I then published as a CSV file. Then, I used the **read_csv** function to put my data into a data frame called **learning_data**. This is the data I will be working with:

```{r}
# EXTRACTING THE DATA AND CREATING OUR DATAFRAME

url <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vRQwki3G6AX9G_3qA-1WBiqXv9wkwXvSLyRsozFDRDu-xAdsimASsZmvAOtt9AAYd1yN926MftmsZrn/pub?gid=147013883&single=true&output=csv"

# rename data variables
learning_data <- read_csv(url) %>%
  rename(time = 1,
       age = 2,
       devices_used = 3,
       most_common_device = 4,
       screen_time = 5,
       most_common_platform = 6,
       job = 7,
       device_job = 8,
       screen_bed = 9)

print(learning_data)
```

First, I discovered that **mobile phones** are the most used devices, primarily used by the 18-25 age group. We can also see the only person to consider an e-reader their most common device was in the **66 and over** age group.

```{r}

# box plot 1: age v.s. most common device

ggplot(data = learning_data) +
  geom_bar(aes(y = most_common_device, fill = age)) +
  labs(title = "Does age change our preference?",
       subtitle = "Comparing the most common device used in different age groups",
       caption = "Source: Natasha Munn 2024",
       x = "Count",
       y = "Device Type") +
  
   theme(panel.background = element_rect(fill = "#F4ECEC"),
    plot.background = element_rect(fill = "#F4ECEC"),
    legend.background = element_rect(fill = "#F4ECEC"),
    legend.box.background = element_rect(fill = "#F4ECEC"))

```

Another aspect of the data I was interested in was the question of screen time before bed, as it is when most people are vulnerable to doom scrolling. I discovered that most respondents are on their screens before bed every night! 

```{r}

# box plot 2: age v.s. days of screen time before bed

ggplot(data = learning_data) +
  geom_bar(aes(x = screen_bed, fill = age)) +
  labs(title = "Who's on their phone the most at bedtime?",
       subtitle = "Comparing the differences in screen use before bed for different ages",
       caption = "Source: Natasha Munn 2024",
       x = "Days of screentime before bed",
       y = "Count") +

 theme(panel.background = element_rect(fill = "#F4ECEC"),
    plot.background = element_rect(fill = "#F4ECEC"),
    axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
    legend.background = element_rect(fill = "#F4ECEC"),
    legend.box.background = element_rect(fill = "#F4ECEC"))
```

The highest amount of total screen time amongst my respondents is *77 hours*. The mean number of days my respondents used a device before bed is *5.7 days*. Here are some more exciting finds from my exploration:

* The most common platform used by my respondents last week was TikTok.

* For the 16 currently employed respondents, it was more common for a device to be essential to their job on a weekly basis.

* The most common device among my respondents in the last week was a mobile phone.

To finish this report, I incorporated some of the **magick packages** I looked at in my last post. Check out this GIF I made celebrating my number of survey results! I used the **paste** function again to input the number of respondents from the learning data into my text. Enjoy!

```{r setup, meme-code}

# meme code

michael <- image_read("https://content.imageresizer.com/images/memes/Michael-Scott-meme-10.jpg")

paste <- paste("Me when I get", length(learning_data$time), "responses on my survey")

text <- image_blank(1043, 100, color = "white") %>%
  image_annotate(paste, size = 30, gravity = "center", font = "Impact") 


#putting the meme together
meme <- c(text, michael) %>%
  image_append(stack = TRUE)

meme

```